---
title: "Esther"
author: "Esther"
format: html
editor: visual
---


## Types and classes
```{r}
typeof(1.5)
typeof("toto")
typeof('also between single quotes')
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```

A type defines a set (math way) with associated operations.


### Classes
```{r}
data(iris)
typeof(iris)
class(iris)
```

Classes is the way to create your own type. Classes is what you create, type is your environement (= what is given to you).
Not much classes. 
2 ways to create classes 
- from a list
- from a vector

## Vectors
Many things in R are vectors (and most of the other things are lists).
A vector is a collection of values of the same type (possibly a single value).

```{r}
x<- 15.25
length(x) ##all vectors have a length
x[1]
```

```{r}
y <- c(1, -2, 45, 1.5, -1e-10)
y 
typeof(y)
y[2]
length(y)
```

```{r}
c(x,y)
```


### Character vectors
This is not R strength...

```{r}
z <- "Alice"
length(z)
```

Salvation comes from the 'stringr' package

```{r}
library(stringr)
```

```{r}
str_length(z)
```

```{r}
complicated <- "🤠"
complicated
length(complicated)
str_length(complicated)
```

### Indexing
```{r}
x <- c(6,5,1,2,3,4)
x[3]
```

```{r}
x[4]<- 15
x
```
Assignement can be understood as a value semantic assignement : 'x' and 'y' are different vectors with the same initial content.


```{r}
y <- x
y
```
Therefore, modifying 'x' has no effect on 'y'.

```{r}
x[1] <- 18
x
y
```
Indexing is made with vectors of any length

```{r}
x
x[c(2, 2, 1, 4, 3)]
```

```{r}
x[0]
x[7]
x[10]
```
```{r}
typeof(numeric(0))
length(numeric(0))
```


```{r}
typeof(NA)
length(NA)
```
```{r}
x
x[c(4, 1, 0, 15)]
```


```{r}
c(numeric(0), 5, numeric(0))
```

```{r}
x
x[c(-1, -4)]
```

```{r}
y <- c("A", "B", "C")
```


## Lists 
vectors must be type uniform, which will trigger a conversion to character in many cases.

```{r}
broken <- c(1L, 1.5, "toto", FALSE)
broken
typeof(broken)
```


Lists do not have limitation

```{r}
bar <- list(1L, 1.5, "toto", FALSE)
bar
```
```{r}
typeof(bar)
length(bar)
```


```{r}
foo <- list(c(1, 2, 3), TRUE, length)
foo
length(foo)
```

### Indexing

```{r}
foo
foo[1]
foo [-2]
foo[c(3, 2, 2, 1)]
```


### Side step : weird values 

```{r}
NA ## missing value
NULL ## missing object
NaN ## Not a number
```

```{r}
0/0
1/0
-1/0
1/0 - 1/0
```

### Back to indexing

```{r}
x<- list(c(1, 2, 3), c(TRUE, FALSE), LETTERS[10:15])
x
```


```{r}
x[c(1, 4, 5)]
```

```{r}
x[5] <- c(-2, 1)
x
```

```{r}
x[5] <- c(-2, 1)
x
```


```{r}
y <- c(1, 2, 3)
y
y[4] <-"toto"
y
```

