---
title: "merged doc"
author: "Rodrigo & Esther"
format: html
editor: visual
---

## Types and classes

```{r}
typeof(1.5)
typeof("toto")
typeof('also between single quotes')
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```

A type defines a set (math way) with associated operations.

### Classes

```{r}
data(iris)
typeof(iris)
class(iris)
```

Classes is the way to create your own type. Classes is what you create, type is your environement (= what is given to you). Not much classes. 2 ways to create classes - from a list - from a vector




## Vectors 

Many things in R are vectors(and most of the other things are lists)

A vector is a collection of values of the same type (possibly a single value)

```{r}
w<- 15.25
length(w)
w[1]

```


```{r}
length(length(w))
typeof(length(w))
```


```{r}
y <- c(1,6,90,-34,1.5,18,1e-12)
y
typeof(y)
y[2]
length(y)
```

c function can be seen as a creation function or, more accurately, as a vector concatenetion function

```{r}
c(w,y)
```


### Character vectors

```{r}
z <- "Alice"
length(z)
```

```{r}
library(stringr)
```

```{r}
str_length(z)
```


```{r}
emojis <- "😀😃🤣😎"
emojis
length(emojis)
str_length(emojis)
```

```{r}
emojis_2 <-c( "😀😃", "🤣😎")
emojis_2
length(emojis_2)
str_length(emojis_2)
