---
title: "Low Level R"
author: "Rodrigo Tellez de Ita"
format: html
---


## Types and classes

### Types

```{r}
typeof(1.5)
typeof("toto")
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```

### Classes

```{r}
data(iris)
typeof(iris)
class(iris)
```

## Vectors 

Many things in R are vectors(and most of the other things are lists)

A vector is a collection of values of the same type (possibly a single value)

```{r}
w<- 15.25
length(w)
w[1]

```


```{r}
length(length(w))
typeof(length(w))
```


```{r}
y <- c(1,6,90,-34,1.5,18,1e-12)
y
typeof(y)
y[2]
length(y)
```

c function can be seen as a creation function or, more accurately, as a vector concatenetion function

```{r}
c(w,y)
```


### Character vectors

```{r}
z <- "Alice"
length(z)
```

```{r}
library(stringr)
```

```{r}
str_length(z)
```


```{r}
emojis <- "😀😃🤣😎"
emojis
length(emojis)
str_length(emojis)
```

```{r}
emojis_2 <-c( "😀😃", "🤣😎")
emojis_2
length(emojis_2)
str_length(emojis_2)
```


## Indexing

```{r}
x <- c(18,6,1,2,3,4)
x[3]

```
```{r}
x[4] <- 15
x
```

Assignment can be understood has a value semantic assignment: 'x' and 'm' are different vectors witht the same initial content.

```{r}
m <- x
m

```


```{r}
x[1] <- 12
x
m
```


Therefore modifying 'x' has no effect on 'm'



Indexing is made with vectors of any length

```{r}
x
x[c(2,2,1,4,3)]
```

```{r}
x[0] ## no possible meaning -> empty numeric vector
x[7] ## no value there -> not avalaible

```

```{r}
typeof(numeric(0))
length(numeric(0))
```

```{r}
typeof(NA)
length(NA)
typeof(NA_integer_)
```


```{r}
x[c(4,1,2,0,12)]
c(numeric(0),5,numeric(0))
```

```{r}
x
x[c(-1,-4)]
```


```{r}
x
x[c(1,4)] <- c(-34,42)
x
x[c(2,10)] <- c(123,78)
x
```


```{r}
a <- c("A","B","C")
a
a[c(TRUE,FALSE,TRUE)]
```



## Lists

```{r}
broken <- c(1L,1.5,"toto",FALSE)
broken
typeof(broken)
```

Vectors must be type uniform, which will trigger a conversion to character in many cases.

Lists do not have this limitation

### Creating a list

```{r}
bar <- list(1L,1.5,"toto",FALSE)
bar
```

```{r}
typeof(bar)
length(bar)
```



```{r}
foo <- list(c(1,2,3),TRUE,length)
foo
length(foo)
```

### Indexing

```{r}
foo
foo[1]
cat("###\n")
foo[-2]
cat("###\n")
foo[c(3,2,2,1)]


```

```{r}
foo[0]
foo[4]
```

```{r}
0/0
1/0
-1/0
1/0-1/0
1/0+1
1/Inf
```
```{r}
NA_real_+1
1/NA_real_
```

```{r}
NULL + 1
```

### Back to indexing lists


The indexing operation '[]' does not change the type or the class of the corresponding object


### Double brackets


```{r}
li <- list(c(1,2,3),c(TRUE,FALSE),LETTERS[10:15])
li[1]
li[[1]]
```

```{r}
li[[2]] <- c("toto","bar")
li[[4]] <- 1:15 # sequence of integer from 1 to 15
li
```

```{r}
li[[c(1,2)]]
li[[c(3,2)]]
```

```{r}
l <- list(1:4,list(letters[1:5],list(c("foo","bar"))))
l
l[[2]]
l[[c(2,1)]]
l[[c(2,1)]]
l[[c(2,2,1,2)]]
l[[2]][[2]][[1]][[2]]
```

### Names

```{r}
n <- list(keys=c('A','B','C'),values=c(10,11,12),key=c(TRUE,FALSE))
n
n[1]
n[[1]]
n$keys 
n$val ## funny but dangerous name compilation
n$key
n$ke
n[["keys"]]
n[["val"]] ## exact matching only
n["values"]
n["values"]|>class()
n[["values"]]|>class()

```

Remark: a data frame is a list in which each element has the same length


```{r}
names(n)
```

```{r}
names(n) <- c("A","B","C")
n

```

## Functions


R extensions (packages) consist essentially in collections of functions

A function is an object that can be called by adding parentheses after the function name, and possibly parameters inside the parameters (effective parameters)


```{r}
names
```
```{r}
log
```

In R functions can have multiple formal parameters and some of them may have default values. When a function is called, one has to specify in general at least the parameters for the parameters without a default value


```{r}
log(10)
log(10,10)
```


```{r}
seq.int
seq.int(1,10)
seq.int(10,1)

```


### Creating functions

This is done with the `function` key word

```{r}
my_fun <- function(x,y) {## between parentheses, we specify the formal ##parameters
  ## between the curcly brackets we desribe the operarions impleteted by the ##function
  
  2*abs(x-y)/(abs(x)+abs(y)+1e-15)
  
}
```


```{r}
my_fun
functionBody(my_fun)
```

```{r}
#| eval: false

my_fun() ## cannot work, x and y missing

```

```{r}
my_fun(2,1)
my_fun(2,2)
my_fun(0,0)
```


Function execution is essentially running the code inside the body of the function after replacing the formal parameters by the values used as effective parameters at the call point. The value of the call is the last value computed in the function.


```{r}
my_fun(4,1)
```

```{r}
my_second_fun <- function(x,y){
  
  z <- x+y
  z/y
  z/x
}
```

```{r}
my_second_fun(1,2)
```

A variable defined in a function is private to the function. It does not care about existing variables with the same name and it disapears just before the end of the function execution.

### Default parameters

To specify a defautlt value for a parameter, we add a equal sign after the formal parameter name followed by default value

```{r}
x_to_y <- function(x,y=2){
  x^y
}
```


```{r}
x_to_y(2,3)
x_to_y(10)
```

Standard calls use positional effective parameters: all effective parameters are specified in the order of the formal parameters.

This can be combined with named effective parameters: the name of the format parameters is used in the call, followed by an equal sign and the chosen value. In this case, order does not matter any more


```{r}
x_to_y(y=4,x=2)
```

Positions and names can be combined if all names are after the unnamed parameters


### Functions are objects

```{r}
all_my_funs <- list(f1=my_fun,f2=my_second_fun,f3=x_to_y)
all_my_funs
```

```{r}
all_my_funs[["f3"]](10,5.5)
```

R has some features of functional languages.

```{r}
listx <- list(1:3,letters[10:20])
listx
```
```{r}
lapply(listx, length)
```
`lapply`applies a function to each of the elements of a list and returns a list made with the results.

This works also on a vector (the result is still a list)

```{r}
lapply(1:10,x_to_y)
```


`sapply`tries to simplify the resulting list into a vector if the content of the list has a single type.

```{r}
sapply(listx, length)
sapply(1:10,x_to_y)
```

```{r}
#| eval: false

res <- lapply(X,f,toto=2,z=32) ##
res[[i]] <- f(X[[i]],toto=2,z=32)
```
Here each call to `x_to_y`has a first effective parameter taken in the vector `1:10`and a second named parameter `y=3`

```{r}
sapply(1:10,x_to_y,y=3)
```

```{r}
my_data <- list(x=1:10,
                y=c(NA,2:5,NA,6:20))

my_data
```
```{r}
lapply(my_data,mean)
```

In this version, the parameter `na.rm` is set to `TRUE` in all the mean calculations



```{r}
lapply(my_data,mean,na.rm=TRUE)
```

### Anonymous functions a.k.a lambdas

Compact function declaration (can be used directly as parameters)

```{r}
\(x) x^2
```

```{r}
log(1:10,base=10)
```

A complicated way of doing sthg simple

```{r}
sapply(1:10, \(x,y) 1+y*log(x),y=2)
1+2*log(1:10)
```

```{r}
lapply(my_data,\(x) mean(x,na.rm = TRUE))
```


