---
title: "Low Level R"
author: "Rodrigo Tellez de Ita"
format: html
---


## Types and classes

### Types

```{r}
typeof(1.5)
typeof("toto")
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```

### Classes

```{r}
data(iris)
typeof(iris)
class(iris)
```

## Vectors 

Many things in R are vectors(and most of the other things are lists)

A vector is a collection of values of the same type (possibly a single value)

```{r}
w<- 15.25
length(w)
w[1]

```


```{r}
length(length(w))
typeof(length(w))
```


```{r}
y <- c(1,6,90,-34,1.5,18,1e-12)
y
typeof(y)
y[2]
length(y)
```

c function can be seen as a creation function or, more accurately, as a vector concatenetion function

```{r}
c(w,y)
```


### Character vectors

```{r}
z <- "Alice"
length(z)
```

```{r}
library(stringr)
```

```{r}
str_length(z)
```


```{r}
emojis <- "😀😃🤣😎"
emojis
length(emojis)
str_length(emojis)
```

```{r}
emojis_2 <-c( "😀😃", "🤣😎")
emojis_2
length(emojis_2)
str_length(emojis_2)
```


## Indexing

```{r}
x <- c(18,6,1,2,3,4)
x[3]

```
```{r}
x[4] <- 15
x
```

Assignment can be understood has a value semantic assignment: 'x' and 'm' are different vectors witht the same initial content.

```{r}
m <- x
m

```


```{r}
x[1] <- 12
x
m
```


Therefore modifying 'x' has no effect on 'm'



Indexing is made with vectors of any length

```{r}
x
x[c(2,2,1,4,3)]
```

```{r}
x[0] ## no possible meaning -> empty numeric vector
x[7] ## no value there -> not avalaible

```

```{r}
typeof(numeric(0))
length(numeric(0))
```

```{r}
typeof(NA)
length(NA)
typeof(NA_integer_)
```


```{r}
x[c(4,1,2,0,12)]
c(numeric(0),5,numeric(0))
```

```{r}
x
x[c(-1,-4)]
```

