---
title: "Low Level R"
author: "Rodrigo Tellez de Ita"
format: html
---


## Types and classes

### Types

```{r}
typeof(1.5)
typeof("toto")
typeof(TRUE)
typeof(1L)
typeof(1)
typeof(typeof)
typeof(list())
```

### Classes

```{r}
data(iris)
typeof(iris)
class(iris)
```

## Vectors 

Many things in R are vectors(and most of the other things are lists)

A vector is a collection of values of the same type (possibly a single value)

```{r}
w<- 15.25
length(w)
w[1]

```


```{r}
length(length(w))
typeof(length(w))
```


```{r}
y <- c(1,6,90,-34,1.5,18,1e-12)
y
typeof(y)
y[2]
length(y)
```

c function can be seen as a creation function or, more accurately, as a vector concatenetion function

```{r}
c(w,y)
```


### Character vectors

```{r}
z <- "Alice"
length(z)
```

```{r}
library(stringr)
```

```{r}
str_length(z)
```


```{r}
emojis <- "😀😃🤣😎"
emojis
length(emojis)
str_length(emojis)
```

```{r}
emojis_2 <-c( "😀😃", "🤣😎")
emojis_2
length(emojis_2)
str_length(emojis_2)
```


## Indexing

```{r}
x <- c(18,6,1,2,3,4)
x[3]

```
```{r}
x[4] <- 15
x
```

Assignment can be understood has a value semantic assignment: 'x' and 'm' are different vectors witht the same initial content.

```{r}
m <- x
m

```


```{r}
x[1] <- 12
x
m
```


Therefore modifying 'x' has no effect on 'm'



Indexing is made with vectors of any length

```{r}
x
x[c(2,2,1,4,3)]
```

```{r}
x[0] ## no possible meaning -> empty numeric vector
x[7] ## no value there -> not avalaible

```

```{r}
typeof(numeric(0))
length(numeric(0))
```

```{r}
typeof(NA)
length(NA)
typeof(NA_integer_)
```


```{r}
x[c(4,1,2,0,12)]
c(numeric(0),5,numeric(0))
```

```{r}
x
x[c(-1,-4)]
```


```{r}
x
x[c(1,4)] <- c(-34,42)
x
x[c(2,10)] <- c(123,78)
x
```


```{r}
a <- c("A","B","C")
a
a[c(TRUE,FALSE,TRUE)]
```



## Lists

```{r}
broken <- c(1L,1.5,"toto",FALSE)
broken
typeof(broken)
```

Vectors must be type uniform, which will trigger a conversion to character in many cases.

Lists do not have this limitation

### Creating a list

```{r}
bar <- list(1L,1.5,"toto",FALSE)
bar
```

```{r}
typeof(bar)
length(bar)
```



```{r}
foo <- list(c(1,2,3),TRUE,length)
foo
length(foo)
```

### Indexing

```{r}
foo
foo[1]
cat("###\n")
foo[-2]
cat("###\n")
foo[c(3,2,2,1)]


```

```{r}
foo[0]
foo[4]
```

```{r}
0/0
1/0
-1/0
1/0-1/0
1/0+1
1/Inf
```
```{r}
NA_real_+1
1/NA_real_
```

```{r}
NULL + 1
```

### Back to indexing lists


The indexing operation '[]' does not change the type or the class of the corresponding object


### Double brackets


```{r}
li <- list(c(1,2,3),c(TRUE,FALSE),LETTERS[10:15])
li[1]
li[[1]]
```

```{r}
li[[2]] <- c("toto","bar")
li[[4]] <- 1:15 # sequence of integer from 1 to 15
li
```

```{r}
li[[c(1,2)]]
li[[c(3,2)]]
```

```{r}
l <- list(1:4,list(letters[1:5],list(c("foo","bar"))))
l
l[[2]]
l[[c(2,1)]]
l[[c(2,1)]]
l[[c(2,2,1,2)]]
l[[2]][[2]][[1]][[2]]
```

### Names

```{r}
n <- list(keys=c('A','B','C'),values=c(10,11,12),key=c(TRUE,FALSE))
n
n[1]
n[[1]]
n$keys 
n$val ## funny but dangerous name compilation
n$key
n$ke
n[["keys"]]
n[["val"]] ## exact matching only
n["values"]
n["values"]|>class()
n[["values"]]|>class()

```

Remark: a data frame is a list in which each element has the same length


```{r}
names(n)
```

```{r}
names(n) <- c("A","B","C")
n

```
